### 2019-07-10 2-Chapters/2-Chapter/chapter2.tex
Vérifications grâce à emccabedj.ucc.asn.au/checktex.html, avant intégration des commentaires d'Emilie sur ma conclusion

Given a measure $\lambda$, that is usually the natural Lebesgue measure on $\R$, an exponential family of probability distributions is defined as the distributions whose density (relative to $\lambda$) can be written as
% \begin{equation}\label{eq:2:exponentialFamily}
$ \Pr_{\eta}(x | \lambda) = h(x) \exp \left( \eta x - A(\eta) \right)$,
% \end{equation}
for a parameter $\eta$ (the canonical parameter),
% and a sufficient statistic that we restrict to be $T(x)=x$,
and a function $h$.
The cumulant function $A(\eta)$ is entirely determined by $\eta$ and $h$,
as $A(\eta) = \log \left( \int h(x) \exp(\eta x) \lambda(\mathrm{d} x) \right)$.
%
The natural parameter space is the set of values of $\eta$ such that this integral $A(\eta)$ is finite,
and usually the literature focuses on regular and minimal exponential families (\ie, when the natural parameter space is an non-empty open set in $\R$).

@@

\paragraph{About Markov models.}
%
% Markov models: we do not present it in this thesis.
Finally, we note that Markov bandit models have also been considered for Cognitive Radio applications \cite{Liu08}.
They are not studied in this thesis, even though we did implemented them in SMPyBandits.
They were introduced in the $1980$s, by Whittle in \cite{Whittle1988} and Anantharam and others in \cite{Anantharam87b}.
A Markov MAB model maps an arm to a Markov chain \cite{Norris98}, instead of a distribution, and thus they are no longer stochastic.
A Markov model exists in two flavors: rested or restless.
For $K$ arms, each Markov chain has a finite number of states $s$, each corresponding to a (constant) reward that the player obtains if she selects this arm while its Markov chain is in state $s$.
Rested Markov models means that only the state of the selected arm's Markov chain can change, following its Markov transition matrix.
Restless models remove this hypothesis, making them harder to track and solve.
%
Such models were less studied than stationary or adversarial models, but some interesting works applied Markov models to Cognitive Radio in the last 10 years
For instance, \cite{Melian15} proposed a CR model mixing MAB and Hidden Markov Models (HMM), solved by a mixed policy called UCB-HMM.
A reader curious about Markov chains could start to read Chapter~3 of \cite{LattimoreBanditAlgorithmsBook} (Section~3.2) and then refer to \cite{Norris98}.

@@

If the goal of the player is to maximize $\E\left[\sum_{t=1}^T r(t)\right]$,
the optimal strategy for this bandit problem is to always pull one of the optimal arms (it can be not unique), but it is unrealistic as the player does not know the true means, nor the optimal arms (as long as they are not all optimal, \ie, in non trivial problems).
%
Comparing the difference between the performance of a fixed baseline and that of the player is a common approach in machine learning research,
and here we can compare with the oracle strategy that always obtains an (expected) reward of $\mu^*$.
%
For a fixed horizon $T$, if $k^*$ denotes the index of any optimal arm,
let us introduce the (mean) regret $R_T^{\cA}$ of an algorithm $\cA$ as
$R_T^{\cA} \eqdef \E\left[ \sum_{t=1}^T (Y_{k^*,t} - r(t)) \right]$.
%
As the rewards from arm $k^*$ are \iid{} (as for all arms), and by linearity of the expectation, we can rewrite this expression to obtain the following definition of the regret.
% \begin{align*}
%     R_T^{\cA}
%     & = \E\left[ \sum_{t=1}^T (Y_{k^*,t} - r(t)) \right]
%     = \E\left[ \sum_{t=1}^T \underbrace{Y_{k^*,t}}_{\text{\iid{} from $\nu_{k^*}$}} \right] - \E\left[ \sum_{t=1}^T r(t) \right] \\
%     & = T \times \E\left[ Y_{k^*,1} \right] - \sum_{t=1}^T \E\left[ r(t) \right]
%     = T \mu^* - \sum_{t=1}^T \E\left[ r(t) \right].
% \end{align*}
% % This is the definition we give below.

@@

\paragraph{Problem-dependent lower-bound in $\Omega(\log T)$.}
\label{par:2:def_optimal_orderoptimal}
%
The following two theorems were proven in \cite{LaiRobbins85}.
These lower-bounds are of highest interest to design efficient algorithms,
and a significant part of the research on MAB algorithms has focused on finding algorithms that matches the Lai and Robbins' lower-bound asymptotically \cite{LaiRobbins85}.
%
This means that a finite-time or an asymptotic upper-bound is proven on the regret of algorithm $\cA$, that asymptotically matches the lower-bound.
When the regret upper-bound matches the lower-bound with the same constant as in the big-$\cO$ notation, we say that the algorithm is asymptotically \emph{optimal},
otherwise the algorithm is said to be \emph{order-optimal} if it matches the lower-bound with with a larger constant.

@@

% - Quantile regret or just histogram of regrets
First, when doing numerical experiments about bandits, if one studies the regret as an empirical mean based on a ``large'' number of random repetitions (\eg, $N=1000$ repetitions), it is important to not only show the mean value but also the variance of the values taken by the regret on each repetition, to verify that all algorithms perform consistently.
Indeed, by only visualizing the mean of $1000$ values, it is possible that we miss some ``bad runs'': if $1$ run out of the $1000$ gives linear regret (\ie, $R_T^{\cA} \propto T$) and the $999$ other give logarithmic regret, then the mean will appear logarithmic.
This is the case of the \Selfish{} algorithm that is defined and explored in Chapter~\ref{chapter:5}.
On the contrary, for efficient algorithms,
we can visualize the entire \emph{distribution} as an histogram, or the variance of the values of $R_T^{\cA}$, and if the number $N$ is reasonably large, we can verify that the regret appears logarithmic for all runs.

@@

% - Worst case regret (max $R_T^{r}$ for $r$ index of Monte-Carlo simulation?)
The \textbf{switching cost} $\mathrm{SC}^{\cA}(T)$ counts how many times the player's decision has changed from one round to the next one, \ie, $\mathrm{SC}^{\cA}(T) \eqdef \sum_{t=1}^{T-1} \indic(A(t) \neq A(t+1))$.
It has recently gained interest in the literature, for instance the authors of \cite{Koren17} proposed an algorithm that adaptively tries to balance the trade-off between minimizing the regret and minimizing the switching cost.
%
Indeed, in single-player models it is easy to show that achieving logarithmic regret directly implies a logarithmic upper-bound on the switching cost, and conversely the lower-bound from \cite{LaiRobbins85} also gives an asymptotic logarithmic lower-bound.
Even if $\mathrm{SC}^{\cA}(T) = \Theta(\log T)$ for an efficient algorithm, it can be interesting to numerically evaluate this quantity, as a large value might indicate an algorithm that is alternating too much between the optimal arm and other arms.
%
It is relevant for cognitive radio applications, as a hardware reconfiguration costs energy, and so changing channel from two consecutive time steps has a non trivial energy cost, as highlighted in \cite{modiDemo2016}.
We also find interesting to study $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, the sum of the switching costs of the $M$ players in a multi-players bandit game (if player $j$ uses algorithm $\cA_j$ for $j=1,\dots,M$), as studied in Chapter~\ref{chapter:5}.
Our proposed algorithm, \MCTopM, achieves order-optimal logarithmic regret as well as a logarithmic number of switches.
% Indeed in this case it is possible that the players attain a logarithmic centralized regret, by following an efficient centralized or decentralized algorithm, while still suffering from a linear switching cost.
% %
% It is very satisfying to obtain a logarithmic switching cost for our algorithm \MCTopM-\klUCB{} in Chapter~\ref{chapter:5}, even if minimizing the switching cost was not one of our objective, and the bound we obtain on $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$ is just a consequence of our proof of the control we give on the regret $R^{\cA}(T)$.
% % - Switching cost? \cite{modiDemo2016} \cite{Koren17}
% On a more experimental note, it was shown in \cite{modiDemo2016} that the previous state-of-the-art policy for multi-players bandits, \rhoRand, could be tuned to run in batches\footnote{~The batch bandit setting implies that all players use the same decisions for instance for $50$ consecutive times, and update their decisions only once every $50$ time steps. For more details, see \cite{modiDemo2016} for experiments on the multi-players case, or \cite{perchet2016,gao2019batched,kolnogorov2019multi} for theoretical developments on the single-player case.}, in order to reduce by a certain multiplicative factor its switching cost while only adding an additive factor on its regret. While such ideas can interest a practitioner, they does not change the asymptotic behavior of $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, which is $\Theta(\log(T))$ for any efficient policy.

@@

-- The uniform strategy always plays the $K$ arms uniformly at random,
$\forall t\in[T], A(t) \sim \cU([K])$, where $\cU(S)$ denotes the uniform distribution on a set $S$, and $\cU([K])$ the uniform distribution on $[K] = \{1,\dots,K\}$ (\ie, $\forall t, \forall k\in[K], \Pr(A(t) = k) = \frac{1}{K}$).
%
The player \textcolor{darkgreen}{\emph{only explores}} without using the collected information, and this strategy fails dramatically for any non trivial problem (\ie, linear regret).
Indeed it obtains a linear (mean) regret $R_T = \frac{1}{K} \sum_{k=1}^K \Delta_k T$
which gives $R_T \propto T$ for any problem with at least one sub-optimal arms (\ie, all non trivial problems, ruling out the case where $\mu_1=\mu_2=\dots=\mu_K$).

@@

\begin{theorem}[Regret bound for $\UCB_1$]\label{thm:2:UCBregretBound}
\begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
    For any instance $I$ with $K$ arms with bounded distributions in $[0,1]$,
    of means $\mu_1,\dots,\mu_k$, and any horizon $T>K$,
    the $\UCB_1$ algorithm with its parameter\footnotemark{} $\alpha=2$ verifies the following finite-time regret bound
    \begin{equation}\label{eq:2:UCBregretBound}
        R_T^{\UCB_1}(I) \leq \left(\sum_{k: \Delta_k>0} \frac{8}{\Delta_k}\right) \log(T) + \sum_{k: \Delta_k>0} \Delta_k \left(1 + \frac{\pi^2}{3}\right).
        % = \bigO{ \left(\sum_{k: \Delta_k>0} \frac{8}{\Delta_k}\right) \log(T)}.
    \end{equation}
    % \vspace*{-20pt}  % FIXME vspace negative to compensate the \end{leftbar}
\end{leftbar}  % XXX leftbar theorembar, comment if needed
\end{theorem}
%
\footnotetext{~For simplicity we present the theorem and its prove in the restricted case of $\alpha=2$. More details on this proof can be found for instance in Section~1.3 of \cite{Slivkins2019}, while for instance Chapter~2 of \cite{Bubeck12} Theorem~2.1) and Chapter~7 of \cite{LattimoreBanditAlgorithmsBook} (Theorem~7.2) both give more general proofs, in particular for any value of $\alpha\geq\frac{1}{2}$.}%
%
% \begin{small}
\begin{smallproof}\label{proof:2:UCBregretBound}
    We focus on bounded distributions in $[0,1]$, for which Hoeffding's inequality gives the inequality \eqref{eq:2:HoeffdingInequality}, as stated above in Lemma~\ref{lem:2:HoeffdingInequality}.
    Thanks to the regret decomposition from Proposition~\ref{prop:2:RegretDecomposition}, we can focus on bounding $\E[N_k(T)]$ for any sub-optimal arm $k$.\\
    %
    \indent
    The first trick consists in writing the following, which is true for any $n \geq 1$:
    \begin{equation}\label{eq:2:trickInUCBProof}
        N_k(T) = 1 + \sum_{t=K+1}^T \indic(A(t) = k)
        \leq 1 + n + \sum_{t=n}^T \indic(A(t) = k, N_k(t) \geq n).
    \end{equation}

@@

Instead of using a confidence radius $\xi_k(t)$ based on Hoeffding's inequality,
the \klUCB{} algorithm rather uses the Chernoff's concentration inequality for the considered exponential family \cite{chernoff1952measure,chernoff1981note}.
% such as Chernoff-Hoeffding for Gaussian (or sub-Gaussian) rewards.
% It yields improved upper confidence bounds (\ie, with smaller $\xi_k(t)$) that are based on Kullback-Leibler divergence.
We remind that we consider rewards bounded in $[0,1]$.
Consider an exploration function $f(t)$,
that is typically chosen as $f(t) \eqdef \log(t)+c\log(\log(t))$ for the theoretical analyses (with $c \geq 3$),
and $f(t) \eqdef \log(t)$ in practice,
then the \klUCB{} indexes are defined by
% The constant $c>0$ is usually chosen as $c = 1$.
\begin{align}\label{eq:2:klUCB_index}
    U_k^{\klUCB}(t) \eqdef \sup\limits_{q \in [0, 1]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{f(t}{N_k(t)} \right\}.
\end{align}

@@

% ----------------------------------------------------------------------------
\textbf{Policies for adversarial bandits.}
%
Another important family of bandits algorithms are the algorithms proposed for the adversarial setting \cite{Auer02NonStochastic}.
The first of them is the \ExpThree{} policy, for ``\emph{Exp}onential weights for \emph{Exp}loitation and \emph{Exp}loration'',
which is an example of the well-known \emph{multiplicative weights update algorithm}\footnote{~For more details, see
% on Wikipedia \href{https://en.wikipedia.org/wiki/Multiplicative_weight_update_method}{\texttt{en.wikipedia.org/wiki/Multiplicative\_weight\_update\_method}},
this article by J. Kun, \href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{JeremyKun.com/2017/02/27/the-reasonable-effectiveness-}}\\
\href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{of-the-multiplicative-weights-update-algorithm}}. The survey \cite{Arora2012multiplicative} gives more details.},
% which comes with an interactive online demonstration to help understanding the behavior of the algorithm, hosted at \href{https://j2kun.github.io/mwua/index.html}{\texttt{j2kun.GitHub.io/mwua/}} (accessed the 24th of May 2019)
a family of algorithms developed since the 1950s, and applied to a wide range of different domains.
% https://smpybandits.github.io/docs/Policies.Exp3.html
%
The \ExpThree{} policy maintains weights $w_k(t)$ on the $K$ arms, and at time $t$ it samples an arm from the distribution which is a mixture of $\bm{w}(t)=[w_1(t),\dots,w_K(t)]$ and the uniform distribution, \ie,
$\Pr(A(t)=k) \eqdef (1-\gamma) w_k(t) + \gamma \frac{1}{K}$.
The weights are initially uniform, \ie, $w_k(0)=1/K$, and then at each time after observing a reward $Y_{k,t}$ from arm $k$, \ExpThree{} updates the weight of the chosen arm multiplicatively, using $w_k(t+1) = w_k(t) \times \exp(\widehat{r_k}(t) / (\gamma N_k(t)))$. The weights are then renormalized, to have a sum $\sum_k w_k(t+1) = 1$.
This notation $\widehat{r_k}(t)$ means an unbiased estimator of the reward $r(t)=Y_{k,t}$, that is, $\widehat{r_k}(t) = r(t) / \Pr(A(t)=k)$.
%
Like for the $\varepsilon$-greedy policy, the parameter $\gamma$ can be constant, or can be a non-increasing sequence $(\gamma_t)_{t\in\N^*}$.
%
It was proven in \cite{Auer02NonStochastic} that \ExpThree{} with constant parameter $\gamma$ achieves $R_T = \bigO{\sqrt{K T \log(T)}}$ problem-independent regret,
while using a decreasing sequence, such as $\gamma_t = \sqrt{\log(K) / K} \log(t) / t$ gives an order-optimal problem-independent regret upper-bound of $\bigO{\sqrt{K T}}$.
A short proof of this results and more details on \ExpThree{} are given in \cite{Bubeck12}.
%
Many other policies have been proposed for adversarial bandits, like Follow-the-Perturbed-Leader or Follow-the-Regularized-Leader for linear adversarial bandits,
% but we do not study them in this thesis,
and we refer to Parts~III and VI of \cite{LattimoreBanditAlgorithmsBook} for more details.

@@

We finish this Section~\ref{sec:2:famousMABalgorithms} by illustrating the performance of different algorithms, on the small MAB problem used in the online demonstration shown above (see Figures~\ref{fig:2:example_of_a_5_arm_bandit_problem} and \ref{fig:2:example_of_a_5_arm_bandit_problem__step100}).
We consider $K=5$ arms following Bernoulli distributions, of means $\mu_1=0.6,\mu_2=0.2,\mu_3=0.55,\mu_4=0.7,\mu_5=0.5$ (unknown to the algorithms), and first an horizon $T=100$ (like in the online demo) then a larger value of $T=10000$.
The best arm is $\mu_4$ and the smallest gap is $\Delta = 0.1$, thus a lower-bound on the gaps is taken as $d=\Delta$.
% (used to tune the two algorithms from Section~\ref{sub:2:naiveSimpleStrategies}).

>>>>

Given a measure $\lambda$, that is usually the natural Lebesgue measure on $\R$, an exponential family of probability distributions is defined as the distributions whose density (relative to $\lambda$) can be written as
% \begin{equation}\label{eq:2:exponentialFamily}
$ \Pr_{\eta}(x | \lambda) = h(x) \exp \left( \eta x - A(\eta) \right)$,
% \end{equation}
for a parameter $\eta$ (the canonical parameter),
% and a sufficient statistic that we restrict to be $T(x)=x$,
and a function $h$.
The cumulant function $A(\eta)$ is entirely determined by $\eta$ and $h$,
as $A(\eta) = \log \left( \int h(x) \exp(\eta x) \lambda(\mathrm{d} x) \right)$.
%
The natural parameter space is the set of values of $\eta$ such that this integral $A(\eta)$ is finite,
and usually the literature focuses on regular and minimal exponential families (\ie, when the natural parameter space is a non-empty open set in $\R$).

@@

\paragraph{About Markov models.}
%
% Markov models: we do not present it in this thesis.
Finally, we note that Markov bandit models have also been considered for Cognitive Radio applications \cite{Liu08}.
They are not studied in this thesis, even though we did implement them in SMPyBandits.
They were introduced in the $1980$s, by Whittle in \cite{Whittle1988} and Anantharam and others in \cite{Anantharam87b}.
A Markov MAB model maps an arm to a Markov chain \cite{Norris98}, instead of a distribution, and thus they are no longer stochastic.
A Markov model exists in two flavors: rested or restless.
For $K$ arms, each Markov chain has a finite number of states $s$, each corresponding to a (constant) reward that the player obtains if she selects this arm while its Markov chain is in state $s$.
Rested Markov models means that only the state of the selected arm's Markov chain can change, following its Markov transition matrix.
Restless models remove this hypothesis, making them harder to track and solve.
%
Such models were less studied than stationary or adversarial models, but some interesting works applied Markov models to Cognitive Radio in the last 10 years
For instance, \cite{Melian15} proposed a CR model mixing MAB and Hidden Markov Models (HMM), solved by a mixed policy called UCB-HMM.
A reader curious about Markov chains could start to read Chapter~3 of \cite{LattimoreBanditAlgorithmsBook} (Section~3.2) and then refer to \cite{Norris98}.

@@

If the goal of the player is to maximize $\E\left[\sum_{t=1}^T r(t)\right]$,
the optimal strategy for this bandit problem is to always pull one of the optimal arms (it can be not unique), but it is unrealistic as the player does not know the true means, nor the optimal arms (as long as they are not all optimal, \ie, in nontrivial problems).
%
Comparing the difference between the performance of a fixed baseline and that of the player is a common approach in machine learning research,
and here we can compare with the oracle strategy that always obtains an (expected) reward of $\mu^*$.
%
For a fixed horizon $T$, if $k^*$ denotes the index of any optimal arm,
let us introduce the (mean) regret $R_T^{\cA}$ of an algorithm $\cA$ as
$R_T^{\cA} \eqdef \E\left[ \sum_{t=1}^T (Y_{k^*,t} - r(t)) \right]$.
%
As the rewards from arm $k^*$ are \iid{} (as for all arms), and by linearity of the expectation, we can rewrite this expression to obtain the following definition of the regret.
% \begin{align*}
%     R_T^{\cA}
%     & = \E\left[ \sum_{t=1}^T (Y_{k^*,t} - r(t)) \right]
%     = \E\left[ \sum_{t=1}^T \underbrace{Y_{k^*,t}}_{\text{\iid{} from $\nu_{k^*}$}} \right] - \E\left[ \sum_{t=1}^T r(t) \right] \\
%     & = T \times \E\left[ Y_{k^*,1} \right] - \sum_{t=1}^T \E\left[ r(t) \right]
%     = T \mu^* - \sum_{t=1}^T \E\left[ r(t) \right].
% \end{align*}
% % This is the definition we give below.

@@

\paragraph{Problem-dependent lower-bound in $\Omega(\log T)$.}
\label{par:2:def_optimal_orderoptimal}
%
The following two theorems were proven in \cite{LaiRobbins85}.
These lower-bounds are of highest interest to design efficient algorithms,
and a significant part of the research on MAB algorithms has focused on finding algorithms that matches the Lai and Robbins' lower-bound asymptotically \cite{LaiRobbins85}.
%
This means that a finite-time or an asymptotic upper-bound is proven on the regret of algorithm $\cA$, that asymptotically matches the lower-bound.
When the regret upper-bound matches the lower-bound with the same constant as in the big-$\cO$ notation, we say that the algorithm is asymptotically \emph{optimal},
otherwise the algorithm is said to be \emph{order-optimal} if it matches the lower-bound with a larger constant.

@@

% - Quantile regret or just histogram of regrets
First, when doing numerical experiments about bandits, if one studies the regret as an empirical mean based on a ``large'' number of random repetitions (\eg, $N=1000$ repetitions), it is important to not only show the mean value but also the variance of the values taken by the regret on each repetition, to verify that all algorithms perform consistently.
Indeed, by only visualizing the mean of $1000$ values, it is possible that we miss some ``bad runs'': if $1$ run out of the $1000$ gives linear regret (\ie, $R_T^{\cA} \propto T$) and the $999$ other give logarithmic regret, then the mean will appear logarithmic.
This is the case of the \Selfish{} algorithm that is defined and explored in Chapter~\ref{chapter:5}.
On the contrary, for efficient algorithms,
we can visualize the entire \emph{distribution} as a histogram, or the variance of the values of $R_T^{\cA}$, and if the number $N$ is reasonably large, we can verify that the regret appears logarithmic for all runs.

@@

% - Worst case regret (max $R_T^{r}$ for $r$ index of Monte-Carlo simulation?)
The \textbf{switching cost} $\mathrm{SC}^{\cA}(T)$ counts how many times the player's decision has changed from one round to the next one, \ie, $\mathrm{SC}^{\cA}(T) \eqdef \sum_{t=1}^{T-1} \indic(A(t) \neq A(t+1))$.
It has recently gained interest in the literature, for instance the authors of \cite{Koren17} proposed an algorithm that adaptively tries to balance the trade-off between minimizing the regret and minimizing the switching cost.
%
Indeed, in single-player models it is easy to show that achieving logarithmic regret directly implies a logarithmic upper-bound on the switching cost, and conversely the lower-bound from \cite{LaiRobbins85} also gives an asymptotic logarithmic lower-bound.
Even if $\mathrm{SC}^{\cA}(T) = \Theta(\log T)$ for an efficient algorithm, it can be interesting to numerically evaluate this quantity, as a large value might indicate an algorithm that is alternating too much between the optimal arm and other arms.
%
It is relevant for cognitive radio applications, as a hardware reconfiguration costs energy, and so changing channel from two consecutive time steps has a nontrivial energy cost, as highlighted in \cite{modiDemo2016}.
We also find interesting to study $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, the sum of the switching costs of the $M$ players in a multi-players bandit game (if player $j$ uses algorithm $\cA_j$ for $j=1,\dots,M$), as studied in Chapter~\ref{chapter:5}.
Our proposed algorithm, \MCTopM, achieves order-optimal logarithmic regret as well as a logarithmic number of switches.
% Indeed in this case it is possible that the players attain a logarithmic centralized regret, by following an efficient centralized or decentralized algorithm, while still suffering from a linear switching cost.
% %
% It is very satisfying to obtain a logarithmic switching cost for our algorithm \MCTopM-\klUCB{} in Chapter~\ref{chapter:5}, even if minimizing the switching cost was not one of our objective, and the bound we obtain on $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$ is just a consequence of our proof of the control we give on the regret $R^{\cA}(T)$.
% % - Switching cost? \cite{modiDemo2016} \cite{Koren17}
% On a more experimental note, it was shown in \cite{modiDemo2016} that the previous state-of-the-art policy for multi-players bandits, \rhoRand, could be tuned to run in batches\footnote{~The batch bandit setting implies that all players use the same decisions for instance for $50$ consecutive times, and update their decisions only once every $50$ time steps. For more details, see \cite{modiDemo2016} for experiments on the multi-players case, or \cite{perchet2016,gao2019batched,kolnogorov2019multi} for theoretical developments on the single-player case.}, in order to reduce by a certain multiplicative factor its switching cost while only adding an additive factor on its regret. While such ideas can interest a practitioner, they does not change the asymptotic behavior of $\mathrm{SC}^{\cA_1,\dots,\cA_M}(T)$, which is $\Theta(\log(T))$ for any efficient policy.

@@

-- The uniform strategy always plays the $K$ arms uniformly at random,
$\forall t\in[T], A(t) \sim \cU([K])$, where $\cU(S)$ denotes the uniform distribution on a set $S$, and $\cU([K])$ the uniform distribution on $[K] = \{1,\dots,K\}$ (\ie, $\forall t, \forall k\in[K], \Pr(A(t) = k) = \frac{1}{K}$).
%
The player \textcolor{darkgreen}{\emph{only explores}} without using the collected information, and this strategy fails dramatically for any nontrivial problem (\ie, linear regret).
Indeed it obtains a linear (mean) regret $R_T = \frac{1}{K} \sum_{k=1}^K \Delta_k T$
which gives $R_T \propto T$ for any problem with at least one sub-optimal arms (\ie, all nontrivial problems, ruling out the case where $\mu_1=\mu_2=\dots=\mu_K$).

@@

\begin{theorem}[Regret bound for $\UCB_1$]\label{thm:2:UCBregretBound}
\begin{leftbar}[theorembar]  % XXX leftbar theorembar, comment if needed
    For any instance $I$ with $K$ arms with bounded distributions in $[0,1]$,
    of means $\mu_1,\dots,\mu_k$, and any horizon $T>K$,
    the $\UCB_1$ algorithm with its parameter\footnotemark{} $\alpha=2$ verifies the following finite-time regret bound
    \begin{equation}\label{eq:2:UCBregretBound}
        R_T^{\UCB_1}(I) \leq \left(\sum_{k: \Delta_k>0} \frac{8}{\Delta_k}\right) \log(T) + \sum_{k: \Delta_k>0} \Delta_k \left(1 + \frac{\pi^2}{3}\right).
        % = \bigO{ \left(\sum_{k: \Delta_k>0} \frac{8}{\Delta_k}\right) \log(T)}.
    \end{equation}
    % \vspace*{-20pt}  % FIXME vspace negative to compensate the \end{leftbar}
\end{leftbar}  % XXX leftbar theorembar, comment if needed
\end{theorem}
%
\footnotetext{~For simplicity we present the theorem and its proof in the restricted case of $\alpha=2$. More details on this proof can be found for instance in Section~1.3 of \cite{Slivkins2019}, while for instance Chapter~2 of \cite{Bubeck12} Theorem~2.1) and Chapter~7 of \cite{LattimoreBanditAlgorithmsBook} (Theorem~7.2) both give more general proofs, in particular for any value of $\alpha\geq\frac{1}{2}$.}%
%
% \begin{small}
\begin{smallproof}\label{proof:2:UCBregretBound}
    We focus on bounded distributions in $[0,1]$, for which Hoeffding's inequality gives the inequality \eqref{eq:2:HoeffdingInequality}, as stated above in Lemma~\ref{lem:2:HoeffdingInequality}.
    Thanks to the regret decomposition from Proposition~\ref{prop:2:RegretDecomposition}, we can focus on bounding $\E[N_k(T)]$ for any sub-optimal arm $k$.\\
    %
    \indent
    The first trick consists in writing the following, which is true for any $n \geq 1$:
    \begin{equation}\label{eq:2:trickInUCBProof}
        N_k(T) = 1 + \sum_{t=K+1}^T \indic(A(t) = k)
        \leq 1 + n + \sum_{t=n}^T \indic(A(t) = k, N_k(t) \geq n).
    \end{equation}

@@

Instead of using a confidence radius $\xi_k(t)$ based on Hoeffding's inequality,
the \klUCB{} algorithm rather uses Chernoff's concentration inequality for the considered exponential family \cite{chernoff1952measure,chernoff1981note}.
% such as Chernoff-Hoeffding for Gaussian (or sub-Gaussian) rewards.
% It yields improved upper confidence bounds (\ie, with smaller $\xi_k(t)$) that are based on Kullback-Leibler divergence.
We remind that we consider rewards bounded in $[0,1]$.
Consider an exploration function $f(t)$,
that is typically chosen as $f(t) \eqdef \log(t)+c\log(\log(t))$ for the theoretical analyses (with $c \geq 3$),
and $f(t) \eqdef \log(t)$ in practice,
then the \klUCB{} indexes are defined by
% The constant $c>0$ is usually chosen as $c = 1$.
\begin{align}\label{eq:2:klUCB_index}
    U_k^{\klUCB}(t) \eqdef \sup\limits_{q \in [0, 1]} \left\{ q : \kl\left(\frac{X_k(t)}{N_k(t)}, q\right) \leq \frac{f(t}{N_k(t)} \right\}.
\end{align}

@@

% ----------------------------------------------------------------------------
\textbf{Policies for adversarial bandits.}
%
Another important family of bandits algorithms are the algorithms proposed for the adversarial setting \cite{Auer02NonStochastic}.
The first of them is the \ExpThree{} policy, for ``\emph{Exp}onential weights for \emph{Exp}loitation and \emph{Exp}loration'',
which is an example of the well-known \emph{multiplicative weights update algorithm}\footnote{~For more details, see
% on Wikipedia \href{https://en.wikipedia.org/wiki/Multiplicative_weight_update_method}{\texttt{en.wikipedia.org/wiki/Multiplicative\_weight\_update\_method}},
this article by J. Kun, \href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{JeremyKun.com/2017/02/27/the-reasonable-effectiveness-}}\\
\href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{of-the-multiplicative-weights-update-algorithm}}. The survey \cite{Arora2012multiplicative} gives more details.},
% which comes with an interactive online demonstration in order to better understand the behavior of the algorithm, hosted at \href{https://j2kun.github.io/mwua/index.html}{\texttt{j2kun.GitHub.io/mwua/}} (accessed the 24th of May 2019)
a family of algorithms developed since the 1950s, and applied to a wide range of different domains.
% https://smpybandits.github.io/docs/Policies.Exp3.html
%
The \ExpThree{} policy maintains weights $w_k(t)$ on the $K$ arms, and at time $t$ it samples an arm from the distribution which is a mixture of $\bm{w}(t)=[w_1(t),\dots,w_K(t)]$ and the uniform distribution, \ie,
$\Pr(A(t)=k) \eqdef (1-\gamma) w_k(t) + \gamma \frac{1}{K}$.
The weights are initially uniform, \ie, $w_k(0)=1/K$, and then at each time after observing a reward $Y_{k,t}$ from arm $k$, \ExpThree{} updates the weight of the chosen arm multiplicatively, using $w_k(t+1) = w_k(t) \times \exp(\widehat{r_k}(t) / (\gamma N_k(t)))$. The weights are then renormalized, to have a sum $\sum_k w_k(t+1) = 1$.
This notation $\widehat{r_k}(t)$ means an unbiased estimator of the reward $r(t)=Y_{k,t}$, that is, $\widehat{r_k}(t) = r(t) / \Pr(A(t)=k)$.
%
Like for the $\varepsilon$-greedy policy, the parameter $\gamma$ can be constant, or can be a non-increasing sequence $(\gamma_t)_{t\in\N^*}$.
%
It was proven in \cite{Auer02NonStochastic} that \ExpThree{} with constant parameter $\gamma$ achieves $R_T = \bigO{\sqrt{K T \log(T)}}$ problem-independent regret,
while using a decreasing sequence, such as $\gamma_t = \sqrt{\log(K) / K} \log(t) / t$ gives an order-optimal problem-independent regret upper-bound of $\bigO{\sqrt{K T}}$.
A short proof of this results and more details on \ExpThree{} are given in \cite{Bubeck12}.
%
Many other policies have been proposed for adversarial bandits, like Follow-the-Perturbed-Leader or Follow-the-Regularized-Leader for linear adversarial bandits,
% but we do not study them in this thesis,
and we refer to Parts~III and VI of \cite{LattimoreBanditAlgorithmsBook} for more details.

@@

We finish this Section~\ref{sec:2:famousMABalgorithms} by illustrating the performance of different algorithms, on the small MAB problem used in the online demonstration shown above (see Figures~\ref{fig:2:example_of_a_5_arm_bandit_problem} and \ref{fig:2:example_of_a_5_arm_bandit_problem__step100}).
We consider $K=5$ arms following Bernoulli distributions, of means $\mu_1=0.6,\mu_2=0.2,\mu_3=0.55,\mu_4=0.7,\mu_5=0.5$ (unknown to the algorithms), and first a horizon $T=100$ (like in the online demo) then a larger value of $T=10000$.
The best arm is $\mu_4$ and the smallest gap is $\Delta = 0.1$, thus a lower-bound on the gaps is taken as $d=\Delta$.
% (used to tune the two algorithms from Section~\ref{sub:2:naiveSimpleStrategies}).

The different algorithms being compared are the following, in the order they were presented above:
%
first we include the pure-exploration and
pure-exploitation algorithms,
then the $\varepsilon$-greedy (with $\varepsilon_t=\varepsilon_0/t$ and $\varepsilon_0=6K/d^2=3000$) and
Explore-then-Commit (with $T_0=K4/d^2 \log(d^2T/4)=2000\log(T/400)$) algorithms
from Section~\ref{sub:2:naiveSimpleStrategies}, which both uses a tuning based on a prior knowledge of $\Delta$ (that gives large values of $\varepsilon_0$ and $T_0$ yielding linear regret for small horizon).
%
We then include efficient algorithms, that do not need prior knowledge of the $T$ (they are anytime) nor $\Delta$ (they are robust and efficient on unknown problems):
$\UCB_1$ (with $\alpha=1$),
\klUCB{} (with Bernoulli $\kl = \KL_{\cB}$),
and Thompson sampling (with Beta posteriors and flat uniform priors).
%
